Complete Algebraic File Synchronization

### 2015-12-19 2015/intro.tex
New introduction

Synchronizing data structures, and filesystems in particular,
is a task that we encounter more and more often in our world
of multiple mobile devices, cloud solutions and distributed systems.
Applications of synchronization ranges from
solutions for end users who expect to access personal data seamlessly
on multiple devices and personal clouds to geographically distributed server networks
that need to synchronize themselves with low latency to serve the 
same content.

In this paper we
continue the work in \cite{NREC} and add to the theoretical understanding
of synchronization based not on filesystem states, but filesystem changes or commands.

% TODO operation-based vs state-based

% Synchronization:
% Main aim: apply updates (where possible) from other replicas
% update detection -> conflict detection -> merging updates -> ordering updates -> apply updates from other replicas
% Problems: conflict detection, ordering

% ORDERING - because when taking updates from different replicas we get an essentially unordered set
% Insight: (Prof. Norman Ramsey)
% - updates where two orders (AB vs BA) are not equivalent are incompatible
% - we only have a set of updates and need to create an order

% New:
% better command (instruction) set
% simpler reconciliation algorithm
% proof of correctness and completeness

% rigorous proof, non-trivial

% Update detection
% - ordering
% Reconciliation
% - conditional operation & inverse sequences

>>>>

Synchronization of data structures
is a mechanism behind many services we take for granted today:
accessing and editing calendar events, documents and spreadsheets
on multiple devices, version control systems and
geographically distributed internet and web services that
guarantee a fast and responsive user experience.

In this paper we investigate an operation- or command-based approach
to filesystem synchronization.
Presented with multiple copies (replicas) of a filesystem that have been independently modified,
we regard the aim of the synchronizer to modify these replicas further
to make them as similar as possible.
The synchronization algorithm we describe
follows the two main steps described by Balasubramaniam and Pierce (\cite{BP}):
update detection, where we identify modifications that have been applied to the replicas
since they diverged,
and reconciliation, where we identify updates that can be propagated to other replicas and do so.
The remaining updates, which could not be propagated, are conflicting updates;
resolving these requires a separate system or human intervention.

In this paper we build on the work done 
on command-based algebraic file synchronization
in \cite{NREC} and add to the theoretical understanding
of synchronization by providing rigorous proofs that the algorithms
we describe work as intended
and offer a solution in our framework that cannot be improved on.

A central problem of command-based synchronisers during both update detection and reconciliation
is ordering.
If update detection is based on comparing the original state of the filesystem
to its new state, then we can easily collect a set of updates,
but we need to order these in a way that they could be applied in
without causing an error 
(for example, a directory needs to be created first before a file can be created under it).
Similarly, during reconciliation when we consider all the updates that
have been applied to the different replicas, we need to find a way of
merging these and creating a global ordering that we can apply
to the filesystems.
In fact, an insight in \cite{NREC} is that
if the ordering changes the effect of two commands,
then under certain circumstances they are not compatible and will give rise to conflicts.
Accordingly, in this paper command-pairs and their properties
play a special role.

In \cref{sec_def} we define a model of filesystems and a set of commands
that we will use to describe updates and modifications.
This is followed by investigating the properties and behaviors
of command pairs in \cref{section_axioms}.
\Cref{sec_update} describes update detection, and it is
in \cref{ordering} that we describe how commands can be ordered
and prove that under some conditions are valid orders
have the same effect.

Finally, \cref{sec_rec} defines our reconciliation algorithm
and when two updates from different replicas give rise to a conflict.
We then proceed to prove that the output of this algorithm
is sound in the sense that 
it can be applied to the replicas without causing an error,
and that the described algorithm provides a complete solution
inasmuch as no further updates can be applied under any circumstances.
In order to be able to do this, we introduce
the concept of {\emph conditional operation} of commands,
and show that it has a number of highly convenient properties.

The approach in the present paper offers an improvement
over rersults presented in \cite{NREC} in multiple ways.
We introduce a new set of commands that is symmetric and
captures more information about the updates.
This has the effect that reasoning about commands
becomes simpler as there are fewer edge cases,
and more powerful as with the additional information
content predictions made by the reasoning are more accurate.
In fact, the new command set not only simplifies
our reconciliation algorithm, but also makes it complete.

The previous work also lacked proofs that the update detection
and reconciliation algorithms it presented work as intended,
which we provide in the current paper.
While the results are intuitive, providing rigorous proofs
is far from trivial.
During the process, we define a number of useful concepts
and show their relationships and the properties they possess.
In our view these construct an algebraic model
that is worthy of interest and further research on its own.
